---
title: "Assignment4: opioid-related deaths"
output: html_notebook
---

Load the tidyverse and janitor libraries:

```{r}

```

Load the data that looks at opioid-related deaths across states and years. Make sure you separate out the notes and clean up the file like we did in class:

```{r}

```

Below each of the following questions, create a code chunk to write code to answer the question. Write your answer in or below the code chunk.

1.  Which year had the most deaths?

2.  Which state had the highest overall death rate?\
    (Note: you cannot sum death rates. You need to sum the deaths and population for each state, then recalculate the rate using the formula we talked about in class.)

3.  How many years of data do we have?

4.  Which states are missing data for one or more years? Does anything in the notes account for this? Does this affect the comparison of overall death rates by state?\
    (Note: one row is one state by one year.)

5.  How does your home state's death rate for 2020 compare with the national death rate?
