---
title: "R Basics"
output:
html_notebook: default
---

This is a chance to practice writing some of the code we've covered so far. 
You can reference our "R Dictionary" here:
[https://github.com/eklucas/ADJS2023/blob/main/tipsheets/R-dictionary.md](https://github.com/eklucas/ADJS2023/blob/main/tipsheets/R-dictionary.md)


Create a *character vector* that lists your five favorite foods (or movies, TV shows, books, albums, whatever) and then print that vector. 
Remember that you have to give it a name.
```{r}

```

Create a *numeric vector* with every whole number between 10 and 20. 
```{r}

```

Pull out the seventh item in that vector.
```{r}

```

Create a data frame with two columns: one for the first names of everyone in your immediate family and one for their ages.
```{r}

```

Pull out the column of last names.
```{r}

```

Pull out the second item in the last names column.
```{r}

```

Import the tidyverse package
```{r}

```

Import the mlb data file
```{r}

```

How many players are pitchers? (note there are right-handed and left-handed pitchers)
```{r}

```

Calculate the pitchers' average salary.
```{r}

```

Which position has the highest average? Which has the highest median? 
```{r}

```

Ask your own questions and answer them: 

Q1: 
```{r}

```

Q2: 
```{r}

```

