---
title: "Cleaning strategies"
output: html_notebook
---

LIBRARIES:

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
#install.packages(readxl) # if needed
library(readxl)
```

DATA (located in Files on Canvas):

```{r}
# deaths file h/t Mary Jo Webster, data editor of the Minnesota Star Tribune
mn_deaths <- read_csv("data/MN_opiate_deaths.csv")

# disney file h/t Andrew Ba Tran, investigative data reporter at the Washington Post
disney <- 

salaries <- 
```

TASKS: `mn_deaths`

1.  Standardize column names

2.  Convert `birthdate`, `deathdate` and `injurydate` to true date fields

3.  Clean up hispanicethnicity

4.  Create an agegroup category

```{r}
# 1
mn_deaths <- mn_deaths %>% clean_names()

# 2
mn_deaths <- mn_deaths %>% mutate(new_birthdate = mdy(birthdate), .after=birthdate)
#note that the .after argument in mutate determines where your new column will go
mn_deaths <- mn_deaths %>% mutate(new_deathdate = mdy(deathdate), .after=deathdate)
mn_deaths <- mn_deaths %>% mutate(new_injurydate = mdy(injury_date), .after=injury_date)

# 3
# note the inconsistencies and issues with `hispanicethnicity`:
mn_deaths %>% count(hispanicethnicity)

# use case_when() to assign values to our `new_ethnicity` column based on certain conditions:
mn_deaths <- mn_deaths %>% 
  mutate(new_ethnicity = case_when(
    str_starts(str_to_lower(hispanicethnicity), "h") ~ "hispanic",
    grepl("not|non", hispanicethnicity, ignore.case=T) ~ "not hispanic",
    TRUE ~ "unknown"), .after=hispanicethnicity)

# 4
# use case_when() to great a new variable `age_groups` based on values in `ageyears`


```


TASKS: `disney`

1.  Convert `release_date` to a true date

2.  Convert `total_gross` and `inflation_adjusted_gross` to numbers fields

TASKS: `salaries`

1.  Convert pay fields to numbers

2.  Clean up department

