---
title: "Cleaning strategies"
output: html_notebook
---

LIBRARIES:

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
#install.packages(readxl) # if needed
library(readxl)
```

DATA (located in Files on Canvas):

```{r}
# deaths file h/t Mary Jo Webster, data editor of the Minnesota Star Tribune
mn_deaths <- read_csv("data/MN_opiate_deaths.csv")

# disney file h/t Andrew Ba Tran, investigative data reporter at the Washington Post
disney <- read_csv("data/disney_movies_total_gross.csv")

# use read_excel() to import an Excel spreadsheet
salaries <- read_excel("data/Bloomington Salaries.xlsx", sheet=1)
```

TASKS: `mn_deaths`

1.  Standardize column names

2.  Convert `birthdate`, `deathdate` and `injurydate` to true date fields

3.  Clean up hispanicethnicity

4.  Create an agegroup category

```{r}
# 1
mn_deaths <- mn_deaths %>% clean_names()

# 2
mn_deaths <- mn_deaths %>% mutate(new_birthdate = mdy(birthdate), .after=birthdate)
#note that the .after argument in mutate determines where your new column will go
mn_deaths <- mn_deaths %>% mutate(new_deathdate = mdy(deathdate), .after=deathdate)
mn_deaths <- mn_deaths %>% mutate(new_injurydate = mdy(injury_date), .after=injury_date)

# 3
# note the inconsistencies and issues with `hispanicethnicity`:
mn_deaths %>% count(hispanicethnicity)

# use case_when() to assign values to our `new_ethnicity` column based on certain conditions:

# in-class solution
mn_deaths <- mn_deaths %>% 
  mutate(new_ethnicity = case_when(
    str_starts(str_to_lower(hispanicethnicity), "h") ~ "hispanic",
    grepl("not|non", hispanicethnicity, ignore.case=T) ~ "not hispanic",
    TRUE ~ "unknown"), .after=hispanicethnicity)

# other solutions
mn_deaths <- mn_deaths %>% mutate(new_ethnicity = case_when(
  hispanicethnicity %in% c("hispanic", "HISPANIC") ~ "hispanic",
  hispanicethnicity %in% c("non-hispanic", "NON-HISPANIC", "not hispanic", "NOT HISPANIC", "NOT-HISPANIC") ~ "non-hispanic", 
  TRUE ~ "unknown"
))

mn_deaths <- mn_deaths %>% mutate(new_ethnicity = case_when(
  str_sub(str_to_lower(hispanicethnicity),1,1) == "h" ~ "hispanic",
  str_sub(str_to_lower(hispanicethnicity),1,1) == "n" ~ "non-hispanic",
  str_sub(str_to_lower(hispanicethnicity),1,1) == "u" ~ "unknown"
))

mn_deaths <- mn_deaths %>% mutate(new_ethnicity = case_when(
  grepl("^h",hispanicethnicity, ignore.case=T) ~ "hispanic",
  grepl("^n", hispanicethnicity, ignore.case=T) ~ "non-hispanic",
  TRUE ~ "unknown"
))

# if you don't use the .after argument of mutate(), you can later use relocate():

mn_deaths <- mn_deaths %>% relocate(new_ethnicity, .after=hispanicethnicity)



# 4
# use case_when() to create a new variable `age_groups` based on values in `ageyears`
# < 18
# 18 - 29
# 30 - 39
# 40 - 49
# 50 - 59
# 60 - 69
# 70 - 79
# 80+

mn_deaths <- mn_deaths %>% mutate(age_groups = case_when(
  ageyears < 18 ~ "under 18",
  ageyears < 30 & ageyears >= 18 ~ "18-29 years",
  ageyears >= 30 & ageyears < 40 ~ "30-39 years",
  ageyears >= 40 & ageyears < 50 ~ "40-49 years",
  ageyears >= 50 & ageyears < 60 ~ "50-59 years",
  ageyears >= 60 & ageyears < 70 ~ "60-69 years",
  ageyears >= 70 & ageyears < 80 ~ "70-79 years",
  ageyears >= 80 ~ "80+ years",
  TRUE ~ "unknown"
))

mn_deaths <- mn_deaths %>% relocate(age_groups, .after=ageyears)

write_csv(mn_deaths, "data/clean_mn_opiate_deaths.csv", na="")

```


TASKS: `disney`

1.  Convert `release_date` to a true date

2.  Convert `total_gross` and `inflation_adjusted_gross` to numbers fields

```{r}
disney <- disney %>% mutate(new_release_dt = mdy(release_date),
                  new_total_gross = parse_number(total_gross),
                  new_adjusted = parse_number(inflation_adjusted_gross))
```

What are the top 10 highest grossing disney movies of all time?
```{r}
disney %>% arrange(desc(new_total_gross))
```

TASKS: `salaries`

1.  Convert pay fields to numbers

2.  Clean up department

```{r}
salaries %>% summarise(gross = sum(total_comp), 
                       overtime = sum(overtime_oncall, na.rm=TRUE),
                       total_sal = sum(salary_2021, na.rm=TRUE))
```

