---
title: "Dispatch Analysis"
output: html_notebook
---

Columbia PD dispatch: https://www.como.gov/CMS/911dispatch/police.php
Downloaded: 1/30/2023
Jan 1, 2017 - Jan 1, 2023

Import libraries & data  
```{r}
library(tidyverse)

dispatch <- read_csv("data/dispatch.csv")
```

80,680 dispatches (one row is one dispatch). 
Has coordinates, need to be converted to a standard system
Things we don't know: 
  - result in arrest? 
  - who was dispatched, or how many
  
Make sure to transform CallDateTime into date type
  

Data vetting: 
Are there duplicate incident numbers? (we don't expect there to be)
```{r}
dispatch %>% group_by(InNum) %>% summarise(n = n()) %>% filter(n > 1)
# there are 42 duplicates

# look at the details to try and figure out why:

dispatch %>% filter(InNum == "2022033169")
# one row just has more information (could delete the other)

dispatch %>% filter(InNum == "2022089607")
# looks like duplicates with different values in ExtNatureDisplayName; want to followup with police on why
```


Run the group_by() and summarise() combo on all your categorical variables. 
The count() function is a shortcut way to do this:

```{r}
# These two lines do the same thing

dispatch %>% group_by(ExtNatureDisplayName) %>% summarise(n = n())

dispatch %>% count(ExtNatureDisplayName) %>% arrange(desc(n))
```

