---
title: "Hot100 Class Questions"
output: html_notebook
---

Libraries
```{r}
library(tidyverse)
library(lubridate)
```

Data
```{r}
hot100 <- read_csv("data/hot-100-current.csv")
```

1. What is the date range for this data? 
```{r}
hot100 %>% summarise(range(chart_week))

```
2. Which songs have been on the chart the longest? 
```{r}
hot100 %>% arrange(desc(wks_on_chart))

hot100 %>% group_by(title, performer) %>% summarise(max = max(wks_on_chart)) %>% arrange(desc(max))
```

3. Which artist had the most #1 songs on the charts? 
```{r}
hot100 %>% 
  filter(current_week==1) %>% 
  group_by(performer) %>% 
  summarise(n = n_distinct(title)) %>% 
  arrange(desc(n))
```

How many #1 songs has Mariah Carey sang in? 
```{r}
hot100 %>% 
  filter(current_week==1) %>% 
  group_by(performer) %>% 
  count(title) %>% 
  filter(grepl("mariah", performer, ignore.case=T))
```

side note: create a table of #1 songs
```{r}
num1_songs <- hot100 %>% 
  filter(current_week==1) %>% 
  distinct(title, performer)

num1_songs %>% count(performer) %>% arrange(desc(n))
```

Looking at Drake
```{r}
hot100 %>% 
  filter(grepl("drake", performer, ignore.case=T) & !performer %in% c("Charlie Drake", "Guy Drake", "Miranda Cosgrove Featuring Drake Bell", "Pete Drake And His Talking Steel Guitar")) %>% 
  count(performer)
```


4. Which song has spent the most weeks at #1? 
```{r}
hot100 %>% 
  filter(current_week==1) %>% 
  count(performer, title) %>% 
  arrange(desc(n))
```

5. Which artist had the most songs in the top 10? 
```{r}
hot100 %>% 
  filter(current_week < 11) %>% 
  distinct(performer, title) %>% 
  count(performer) %>% 
  arrange(desc(n))
```

6. How many times has a song with the word "Christmas" been on the charts?
Which month features the most christmas songs? 
```{r}
hot100 %>% 
  filter(grepl("christmas", title, ignore.case=T)) %>% 
  group_by(month(chart_week)) %>% 
  summarise(n = n_distinct(performer,title))
```

7. How many times has the word "love" appeared in top 100 songs in February?
How does this compare to the other months? 
```{r}
hot100 %>% 
  filter(grepl("love", title, ignore.case=T)) %>% 
  group_by(month(chart_week)) %>% 
  summarise(n = n_distinct(performer,title)) %>% 
  arrange(desc(n))

hot100 %>% 
  filter(grepl("love", title, ignore.case=T)) %>% 
  count(title)
```

7. What was the #1 song the day you were born? 
```{r}
hot100 %>% 
  filter(chart_week <= "1983-05-27") %>% 
  arrange(desc(chart_week), current_week)
```

Which song jumped the highest in one week? 
```{r}
hot100 %>% 
  mutate(jump = case_when(
    is.na(last_week) ~ 101 - current_week,
    last_week == 0 ~ 101 - current_week,
    TRUE ~ last_week - current_week)) %>% 
  arrange(desc(jump))
```

